Measuring the Type Ia supernova delay time distribution via host galaxy rates in the Dark Energy Survey

### 2021-01-19 draft.tex

\begin{table}
	\centering
	\caption{Numbers of field galaxies and SN hosts passing various quality cuts.}
	\label{tab:cuts}
	\begin{tabular}{lccr} % four columns, alignment for each
		\hline
		Cut & N (field galaxies)  & N (SN hosts)\\
		\hline
		Chosen fields & 1364311  & 1766\\
	    Kron mag in all bands & 1069004  & 1764 \\
	    Masked chips & 816950  & - \\
	    Has ugrizJHK photometry & 545748  & -\\
	    Edge of chip & 481731 & 1716 \\
	    Star/galaxy separation & 400051 &  1678\\
	    Has good redshift & 395034 & 1678 \\
	    SNR $\geq 3$& 338256  & 1678 \\
	    $m_r \leq 26.5$ & 196109 &  1678 \\
	    $0.2 \leq z \leq 0.8$ & 48177 & 1414\\

@@

\subsection{Galaxy properties \label{subsec:properties}}

@@

The distribution of efficiency as a function of redshift is shown in Figure \ref{fig:efficiencies}. It is evident that the deep fields (X3, C3) are sensitive to SNe at higher redshifts, while there is no drastic shifts between efficiencies in the eight shallow fields.

@@

To correct for incompleteness caused by the magnitude limited nature of the survey we follow the prescription of e.g. \citet{Sullivan2006, Smith2012} by using a $V_{\mathrm{max}}$ method of \citet{Schmid1968}. For each galaxy in the sample we calculate the maximum volume within which it would have been observed given its absolute magnitude and $k-$correction. A correction of $V_{\mathrm{survey}}/V_{\mathrm{max}}$ is applied to all galaxies for which $V_{\mathrm{max}} < V_{\mathrm{survey}}$, where $V_{\mathrm{survey}}$ is the maximum volume reached by the survey. In our case this corresponds to the volume at $z=0.8$. Figs. \ref{fig:vmax_gals} and \ref{fig:vmax_hist} show the distribution of the correction among galaxies in the sample. The vast majority of galaxies require no correction, meaning they would have been observed beyond the maximum volume considered. Roughly 1\% of objects have a correction greater than 1, with the distribution well described by a power-law function. In Section \ref{sec:rates} we split the samples into redshift bins in order to measure cosmic evolution of the SN Ia rate. For each redshift bin, we reevaluate the value of $V_{\mathrm{survey}}$ used in the incompleteness correction to be the volume of the bin.

@@

The rate of SNe Ia per galaxy as a function of stellar mass is shown in Fig. \ref{fig:rate_raw}. The relationship between SN Ia rate and galaxy stellar mass is well described by a linear function in log space, which corresponds to a power law. To find the slope and intercept that best describe the data we employ Hamiltonian Monte-Carlo (HMC), which we implement via the \texttt{pystan} package and describe in full detail in Appendix \ref{appendix:linear_fits}. We measure a slope of $0.70\pm0.02$, which is broadly consistent with the values found in SNLS and SDSS by \citet{Sullivan2006} and \citet{Smith2012} respectively. It is evident that the linear fit is not consistent with the data across the full range of stellar mass considered. In particular, there appear to be two locations of interest: firstly, the slope is steeper at masses lower than $10^9$ M$_{\odot}$; secondly there is an apparent flattening and possibly a turnover at masses around and above $10^{11}$ M$_{\odot}$. The most likely cause of these divergences is because the SN rate is driven by the SFH which is not a linear function of stellar mass (see Section \ref{sec:method}). Second order effects may also be important, and we address these in the following sections.

>>>>

\begin{table}
	\centering
	\caption{Numbers of field galaxies and SN hosts passing various quality cuts.}
	\label{tab:cuts}
	\begin{tabular}{lccr} % four columns, alignment for each
		\hline
		Cut & N (field galaxies)  & N (SN hosts)\\
		\hline
		Chosen fields & 1364311  & 1766\\
	    Kron mag in all bands & 1069004  & 1764 \\
	    Masked chips & 816950  & - \\
	    Has ugrizJHK photometry & 545748  & -\\
	    Edge of chip & 481731 & 1716 \\
	    Star/galaxy separation & 400051 &  1678\\
	    Has good redshift & 395034 & 1678 \\
	    SNR $\geq 3$& 338256  & 1678 \\
	    $m_r \leq 24.5$ & 196109 &  1678 \\
	    $0.2 \leq z \leq 0.8$ & 48177 & 1414\\

@@

\begin{figure}
    \centering
    \includegraphics[width=0.5\textwidth]{figs/field_vmax.png}
    \caption{Distribution of $V_{\mathrm{max}}$ corrections applied to field galaxies in the DES sample. The majority of galaxies have no correction, and the distribution of the 1\% that do follows a power-law.
    \label{fig:vmax_field}}
\end{figure}
\subsection{Galaxy properties \label{subsec:properties}}

@@

The distribution of efficiency as a function of redshift is shown in Figure \ref{fig:SN_efficiency}. It is evident that the deep fields (X3, C3) are sensitive to SNe at higher redshifts, while there is no drastic shifts between efficiencies in the eight shallow fields.

@@

To correct for incompleteness caused by the magnitude limited nature of the survey we follow the prescription of e.g. \citet{Sullivan2006, Smith2012} by using a $V_{\mathrm{max}}$ method of \citet{Schmidt1968}. For each galaxy in the sample we calculate the maximum volume within which it would have been observed given its absolute magnitude and $k-$correction. A correction of $V_{\mathrm{survey}}/V_{\mathrm{max}}$ is applied to all galaxies for which $V_{\mathrm{max}} < V_{\mathrm{survey}}$, where $V_{\mathrm{survey}}$ is the maximum volume reached by the survey. In our case this corresponds to the volume at $z=0.8$. Fig \ref{fig:vmax_field} shows the distribution of the correction among galaxies in the sample. The vast majority of galaxies require no correction, meaning they would have been observed beyond the maximum volume considered. Roughly 1\% of objects have a correction greater than 1, with the distribution well described by a power-law function. %In Section \ref{sec:rates} we split the samples into redshift bins in order to measure cosmic evolution of the SN Ia rate. For each redshift bin, we reevaluate the value of $V_{\mathrm{survey}}$ used in the incompleteness correction to be the volume of the bin.

@@

The rate of SNe Ia per galaxy as a function of stellar mass is shown in Fig. \ref{fig:rate_raw}. The relationship between SN Ia rate and galaxy stellar mass is well described by a linear function in log space, which corresponds to a power law. To find the slope and intercept that best describe the data we employ Hamiltonian Monte-Carlo (HMC), which we implement via the \texttt{pystan} package and describe in full detail in Appendix \ref{appendix:linear_fits}. We measure a slope of $0.70\pm0.02$, which is broadly consistent with the values found in SNLS and SDSS by \citet{Sullivan2006} and \citet{Smith2012} respectively. It is evident that the linear fit is not consistent with the data across the full range of stellar mass considered. In particular, there appear to be two locations of interest: firstly, the slope is steeper at masses lower than $10^9$ M$_{\odot}$; secondly there is an apparent flattening and possibly a turnover at masses around and above $10^{11}$ M$_{\odot}$. The most likely cause of these divergences is because the SN rate is driven by the SFH which is not a linear function of stellar mass (see Section \ref{sec:model}). Second order effects may also be important, and we address these in the following sections.
